---
title: "Clustering Lab"
output: html_document
date: "2023-11-28"
---

1 -- Load Data
```{r}
univ <- read.csv("Universities.csv", header = TRUE)

# summary of univ
summary(univ)
```

2 -- Remove Records with missing data 
```{r}
univ_complete <- na.omit(univ)
```

3 -- Hirearchial clustering using normalized data, complete linkage and Manhattan distance

3.1 ----- Remove categorical variables and ID
```{r}
univ_cont <- univ_complete[, -c(1:3)]
```

3.2 ----- Normalize data 
```{r}
univ_cont_norm <- sapply(univ_cont, scale)

```

3.3 ----- Compute distance matrix
```{r}
distance_matrix <- dist(univ_cont_norm, method = "manhattan")
head(distance_matrix)
```

3.4 ----- Create hirearchial clusters
```{r}
hierarchical_cluster <- hclust(distance_matrix, method = "complete")

# plot dendrogram
plot(hierarchical_cluster, hang = "-200", ann = TRUE)
```
Check cluster quality
```{r}
library(factoextra)

fviz_nbclust(univ_cont_norm, 
             hcut, method = "silhouette") +
  labs(subtitle = "Silhouette method")
```
3.5 ------ Compute cluster membership, use k = 2
```{r}
univ_hc_m_memb <- cutree(hierarchical_cluster, k = 2)
head(univ_hc_m_memb)
```

3.6 ------- Create cluster dataframe
```{r}
univ_hc_m_memb_df <- as.data.frame(univ_hc_m_memb)
```

Rename cluster
```{r}
names(univ_hc_m_memb_df)[1] <- "Cluster"
```

3.7 ------ Merge cluster memberships with original data
```{r}
univ_w_cluster <- cbind(univ_cont, 
                                     univ_hc_m_memb_df)

head(univ_w_cluster) 
```




