---
title: "Lab_Regression"
output: html_document
date: "2023-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1 --------------------
Load Data
```{r}
airfares <- read.csv("Airfares.csv", header = TRUE)
head(airfares, 10)
# Remove unnecessary variables
# Since S_Code, S_CITY, E_CODE, and E_CITY are likely to determine
# DISTANCE, and these 4 categorical variables have too many values,
# resulting in too many dummy variables,
# they are dropped from the regression model.
# Incidentally, they are also not included in the new record.

airfares <- airfares[, -c(1:4)]
names(airfares)
```

2 --------------------
Training-Validation split and correlations
```{r}
set.seed(666)

train_index <- sample(1:nrow(airfares), 0.6 * nrow(airfares))
valid_index <- setdiff(1:nrow(airfares), train_index)

train_df_st <- airfares[train_index, ]
valid_df_st <- airfares[valid_index, ]

nrow(train_df_st)
nrow(valid_df_st)
```

Correlation matrix 
```{r}
cor(train_df_st[sapply(train_df_st,is.numeric)]) 
```
Distance seems to be the best predictor of FARE, at 0.68 


3 --------------------
Create a model based on the best predictor & evaluate the accuracy
```{r}
Fare_Model <- lm(FARE ~ DISTANCE + COUPON + NEW + HI + S_INCOME + E_INCOME + S_POP + E_POP + PAX,
                      data = train_df_st)
summary(Fare_Model)
```
- Residual shows the model to fit the data well, as it's somewhat spread around 0

- The intercept and most variables are statistically significant, showing 99% + 
confidence in prediction 

- The residual standard error (42.3) is much smaller than the difference between
the min and max of FARE in both the training & validation set (~340), and also
less than the standard deviation (~75) showing the model to be a good fit

- The Adjusted R-squared shows that the model can predict 68.4% of the variance
in FARE

- The F-statistic shows the model is statistically significant


Evaluate accuracy of model on training & validation sets:

```{r}
library(forecast)

# training set
Fare_Model_pred_train <- predict(Fare_Model,
                                train_df_st)

accuracy(Fare_Model_pred_train, train_df_st$FARE)

# validation set
Fare_Model_pred_valid <- predict(Fare_Model,
                                valid_df_st)

accuracy(Fare_Model_pred_valid, valid_df_st$FARE)
```
- The RMSE doesnt vary much from training to validation (41.8 to 42.6)

- There is an average percentage difference of 25.3% between predicted and 
actual values on the validation set, consistent with the 26% difference on training

```{r}
max(train_df_st$FARE) - min(train_df_st$FARE)

sd(train_df_st$FARE)

max(valid_df_st$FARE) - min(valid_df_st$FARE)

sd(valid_df_st$FARE)
```
The variability and range of values are slightly higher in training to 
validation, but overall similar 

4 --------------------
Now we'll create a new record to represent a new route
```{r}
new_record <- data.frame(COUPON = 1.202, NEW = 3, 
                         VACATION = "Yes", SW = "Yes", 
                         HI = 4442.141, S_INCOME = 28760, 
                         E_INCOME = 27664, S_POP = 4557004, 
                         E_POP = 3195503, SLOT = "Free", 
                         GATE = "Free", PAX = 12782, 
                         DISTANCE = 1976)
```

And predict the fare 
```{r}
predicted_fare <- predict(Fare_Model, newdata = new_record)

# Display the predicted fare
cat("Predicted FARE:", predicted_fare, "\n")
```
The predicted Fare for the new record is $234.09

5 ---------------------
The variables likely to actually be available in real world implementation 
of such a model would be variables that don't change: Distance, SW, VACATION, HI, Origin and Destination Income & Population. It's possible that other variables have difficulty in obtaining such data. 

6 ---------------------

6a. A technique I would use to predict whether a route's fare is high or low is
logistic regression. 

b. We would likely need to encode categorical variables into numerical variables,
as well as potentially transform variables for a combined effect.

c. Logistic regression serves a good technique for predicting binary outcomes
such as Low/High. It can also give us coefficients that indicate how each factor 
influences the likelihood of a route having high or low fare.

d. To evaluate the quality of this new model we would look at a confusion matrix
to see how the validation performs on predicting high/low values as well as an
ROC curve to assess it's ability to distinguish true positives.
