---
title: "kNN Lab"
output:
  html_document:
    df_print: paged
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries and data
```{r}
# caret library
library(caret)

# load bank data
bank <- read.csv("UniversalBank.csv")
names(bank)
head(bank,10)
str(bank)
```
Clean up 
```{r}
# drop ID and zip code columns
bank <- bank[ , -c(1, 5)]
names(bank)

# reorder variables. Put the response last
bank <- bank[ , c(1:7, 9:12, 8)]

head(bank)
```

```{r}
# set categorical variables as factor
bank$Education <- as.factor(bank$Education)
bank$Securities.Account <- as.factor(bank$Securities.Account)
bank$CD.Account <- as.factor(bank$CD.Account) 
bank$Online <- as.factor(bank$Online) 
bank$CreditCard <- as.factor(bank$CreditCard) 

# rename outcome variable values
bank$Personal.Loan <- factor(bank$Personal.Loan,
                              levels = c("0", "1"),
                              labels = c("No", "Yes"))

table(bank$Personal.Loan)
```

Set training and validation sets 
```{r}
set.seed(666)

train_index <- sample(1:nrow(bank), 0.6 * nrow(bank))
valid_index <- setdiff(1:nrow(bank), train_index)

train <- bank[train_index, ]
valid <- bank[valid_index, ]

nrow(train)
nrow(valid)
```

Define new customer 
```{r}
new_cust <- data.frame(Age = 40,
                       Experience = 10,
                       Income = 84,
                       Family = 2,
                       CCAvg = 2,
                       Education = 2,
                       Mortgage = 0,
                       Securities.Account = 0,
                       CD.Account = 0,
                       Online = 1,
                       CreditCard = 1)

# set categorical variables as factor

new_cust$Education <- as.factor(new_cust$Education)
new_cust$Securities.Account <- as.factor(new_cust$Securities.Account)
new_cust$CD.Account <- as.factor(new_cust$CD.Account) 
new_cust$Online <- as.factor(new_cust$Online) 
new_cust$CreditCard <- as.factor(new_cust$CreditCard) 

new_cust
```

Prepare for kNN
```{r}
# normalisation for numerical variables

norm_values <- preProcess(train[, -c(6, 8:12)],
                          method = c("center", "scale"))

```
```{r}
train_norm <- train 
valid_norm <- valid
```

```{r}
# normalize the training set

train_norm[, -c(6, 8:12)] <- predict(norm_values,
                                train[, -c(6, 8:12)])
head(train_norm)
```

```{r}
# normalize the validation set 

valid_norm[, -c(6, 8:12)] <- predict(norm_values,
                                     valid[, -c(6, 8:12)])

head(valid_norm)
```

```{r}
# predict normalized values for new customer 

new_cust_norm <- predict(norm_values, new_cust)
new_cust_norm
```

Predict training set across k values 
```{r}
# Train k = 3
knn_model_k3 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 3)
knn_model_k3
```
```{r}
# Predict training set using k = 3

knn_pred_k3_train <- predict(knn_model_k3, 
                             newdata = train_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k3_train)
```
```{r}
# evaluate k = 3
confusionMatrix(knn_pred_k3_train, as.factor(train_norm[, 12]),
                positive = "Yes")
```
k = 5
```{r}
# train k = 5

knn_model_k5 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 5)
knn_model_k5
```
```{r}
# predict training set using k = 5

knn_pred_k5_train <- predict(knn_model_k5, 
                             newdata = train_norm[, -c(12)], 
                             type = "class")

head(knn_pred_k5_train)
```

```{r}
# evaluate k = 5
confusionMatrix(knn_pred_k5_train, as.factor(train_norm[, 12]),
                positive = "Yes")
```
k = 7
```{r}
# train k = 7
knn_model_k7 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 7)
knn_model_k7
```
```{r}
# predict training set using k = 7

knn_pred_k7_train <- predict(knn_model_k7, 
                             newdata = train_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k7_train)
```

```{r}
# evaluate k = 7
confusionMatrix(knn_pred_k7_train, as.factor(train_norm[, 12]),
                positive = "Yes")
```
Predict the validation set across values of k 

k = 3 validation 
```{r}
# predict training set for k = 3
knn_pred_k3_valid <- predict(knn_model_k3, 
                             newdata = valid_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k3_valid)
```

```{r}
# evaluation of k = 3
confusionMatrix(knn_pred_k3_valid, as.factor(valid_norm[, 12]),
                positive = "Yes")
```

```{r}
# load ROSE library
library(ROSE)

# ROC curve for k = 3
ROSE::roc.curve(valid_norm$Personal.Loan, 
                knn_pred_k3_valid)
```

k = 5 validation 
```{r}
# predict training set for k = 5
knn_pred_k5_valid <- predict(knn_model_k5, 
                             newdata = valid_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k5_valid)
```


```{r}
# evaluation of k = 5
confusionMatrix(knn_pred_k5_valid, as.factor(valid_norm[, 12]),
                positive = "Yes")
```

```{r}
# ROC curve for k = 5
ROSE::roc.curve(valid_norm$Personal.Loan, 
                knn_pred_k5_valid)
```

k = 7 validation 
```{r}
# predict training set for k = 7
knn_pred_k7_valid <- predict(knn_model_k7, 
                             newdata = valid_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k7_valid)
```

```{r}
# evaluation of k = 7
confusionMatrix(knn_pred_k7_valid, as.factor(valid_norm[, 12]),
                positive = "Yes")
```

```{r}
# ROC curve for k = 7 
ROSE::roc.curve(valid_norm$Personal.Loan, 
                knn_pred_k7_valid)
```
k = 3 is the best choice for k, it was the most accurate on the training (97.5%) AND the validation set (96.15%).

The AUC of 78.7 proves it to also be the best model, and a good one. 

So we'll use it to predict the new customer:

```{r}
# predict outcome
new_cust_k3_pred <- predict(knn_model_k3, 
                                newdata = new_cust_norm,
                                type = "class")
new_cust_k3_pred
```


```{r}
# predict probability, this shows that all 3 of the nearest neighbors were 'No' 
new_cust_prob_k3 <- predict(knn_model_k3, 
                                newdata = new_cust_norm,
                                type = "prob")
new_cust_prob_k3
```

The new customer is predicted to not accept a loan offer.
